Leadership across cultures and contexts: an exploratory analysis

### 2020-04-30 leadership_across_cultures_contexts.Rmd
reorder leader-follower.
Minor edit.

<!-- INTRO TO MOTIVATE PAPER -->
Research on leadership within anthropology dates to the inception of the discipline. Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917]. Anthropologists documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017]. For review, see @garfield_review_2019.

@@

Since the seminal WEIRD people paper, many social scientists have indeed taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that

>culture is important to anthropology because the anthropological distinction between self and other rests on it. Culture is the essential tool for making other. As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference,anthropology also helps construct, produce, and maintain it. Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages. Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.

>>>>

<!-- INTRO TO MOTIVATE PAPER -->
Research on leadership within anthropology dates to the inception of the discipline. Early on, anthropologists (and some sociologists) realized that leadership and followership were critical to understanding human psychology, social organization, and culture [e.g., @mumford_origins_1906; @firth1927economic; @morgan_ancient_1877; @myres_influence_1917]. Anthropologists have documented some sort of leadership among every ethnographically studied culture [@brown_human_1991; @lewis_leaders_1974], and in many social contexts, such as families and kin groups (Dussart, 2000), ritual [@singh_cultural_2017], work groups [@macfarlan_competitive_2012], and conflicts between groups [@glowacki_evolutionary_2017]. For review, see @garfield_review_2019.

@@

Since the seminal WEIRD people paper, many social scientists have taken up the challenge to diversify their samples and have sought out "non-WEIRD" participants and databases. Dichotomizing populations and their psychologies and social dynamics as WEIRD or non-WEIRD, however, is a mistake.^[A Google Scholar search for "non-WEIRD" produced over 1000 results.] As anthropology has conclusively demonstrated, there is enormous diversity across societies that does not remotely resemble a dichotomy.

Anthropology, on the other hand, has been criticized for exoticizing and essentializing the "Other" [@said1979; @abu2008]. @abu2008 [p. 470], for instance, notes that

>culture is important to anthropology because the anthropological distinction between self and other rests on it. Culture is the essential tool for making other. As a professional discourse that elaborates on the meaning of culture in order to account for, explain, and understand cultural difference, anthropology also helps construct, produce, and maintain it. Anthropological discourse gives cultural difference (and the separation between groups of people it implies) the air of the self-evident. In this regard, the concept of culture operates much like its predecessor - race - even though in its twentieth-century form it has some important political advantages. Unlike race, and unlike even the nineteenth-century sense of culture as a synonym for civilization (contrasted to barbarism), the current concept allows for multiple rather than binary differences.

There is as much diversity within non-Western cultures as there is within Western ones, and there is also substantial overlap of the behaviors of people everywhere [@brown_human_1991].
